### 2022-03-04 ehoh/ehoh.tex
more

\begin{rep}The TPTP includes benchmarks from various areas of
computer science and mathematics. It is the de facto standard for evaluating
automatic provers, but it has few higher-order problems.
%
\end{rep}%
For the first group of benchmarks, we randomly selected 1000 FOF problems
(out of 8172)
and all monomorphic TFF problems that are parsable by E\confrep{}{ within 60\,s
(amounting to 147 out of 231 monomorphic TFF problems)}.
%
Both groups of Sledgehammer problems include two subgroups of 2506 problems,
generated to include 32 or 512 Isabelle lemmas (SH32 and SH512), to
represent both small and large problems.
% arising in interactive
%verification.
Each subgroup consists of two sub-sub\-groups of 1253
problems, generated by using either $\lambda$-lifting or SKBCI-style combinators
to encode $\lambda$-expressions.

\looseness=-1
To ascertain the effectiveness of our approach, we evaluated Ehoh against
E used on applicative encodings of problems (denoted by \appE{}). For reference,
we also evaluated the following versions of higher-order provers that competed in the
THF division of the 2019 edition of CASC \cite{gs-19-casc27}:\
CVC4
1.8 prerelease \cite{brotb-19-ho-smt}, Leo-III 1.4 \cite{sb-21-leo3},
Satallax 3.4 \cite{cb-12-satallax}, Vampire 4.4 \cite{br-19-restricted-unif},
and Zipperposition 1.6 \cite{bbtvw-21-sup-lam}.
Like at CASC, we used different versions of Vampire for first-order
and higher-order problems. Similarly, Zipperposition does not use E as backend when
it is run on first-order problems
and uses different heuristics on first-~and higher-order problems.
The genuine higher-order provers have
the unfair advantage that they can instantiate higher-order
variables with $\lambda$-terms. % (or equivalent SK-style combinator terms).
Thus, some formulas that are
provable by these systems may be nontheorems for \appE{} and Ehoh\confrep{}{, or
they may require tedious reasoning about $\lambda$-lifted functions or SKBCI-style
combinators}.
An example is the conjecture $\exists f.\>\forall x\> y.\; f\>x\;y \eq
\cst{g}\;y\;x$, whose proof requires taking $\lambda x\> y.\; \cst{g}\;y\;x$
as the witness for~$f$.

>>>>

\begin{rep}The TPTP includes benchmarks from various areas of
computer science and mathematics. It is the de facto standard for evaluating
automatic provers, but it has few higher-order problems.
%
\end{rep}%
For the first group of benchmarks, we randomly selected 1000 FOF problems
(out of 8172)
and all monomorphic TFF problems that are parsable by E\confrep{}{ within 60\,s
(amounting to 147 out of 231 monomorphic TFF problems)}.
%
Both groups of Sledgehammer problems include two subgroups of 2506 problems,
generated to include 32 or 512 Isabelle lemmas (SH32 and SH512), to
represent both small and large problems.
% arising in interactive
%verification.
Each subgroup consists of two sub-sub\-groups of 1253
problems, generated by using either $\lambda$-lifting or {\sf SKBCI}-style combinators
to encode $\lambda$-expressions.

\looseness=-1
To ascertain the effectiveness of our approach, we evaluated Ehoh against
E used on applicative encodings of problems (denoted by \appE{}). For reference,
we also evaluated the following versions of higher-order provers that competed in the
THF division of the 2019 edition of CASC \cite{gs-19-casc27}:\
CVC4
1.8 prerelease \cite{brotb-19-ho-smt}, Leo-III 1.4 \cite{sb-21-leo3},
Satallax 3.4 \cite{cb-12-satallax}, Vampire 4.4 \cite{br-19-restricted-unif},
and Zipperposition 1.6 \cite{bbtvw-21-sup-lam}.
Like at CASC, we used different versions of Vampire for first-order
and higher-order problems. Similarly, Zipperposition does not use E as backend when
it is run on first-order problems
and uses different heuristics on first-~and higher-order problems.
The genuine higher-order provers have
the unfair advantage that they can instantiate higher-order
variables with $\lambda$-terms. % (or equivalent SK-style combinator terms).
Thus, some formulas that are
provable by these systems may be nontheorems for \appE{} and Ehoh\confrep{}{, or
they may require tedious reasoning about $\lambda$-lifted functions or {\sf SKBCI}-style
combinators}.
An example is the conjecture $\exists f.\>\forall x\> y.\; f\>x\;y \eq
\cst{g}\;y\;x$, whose proof requires taking $\lambda x\> y.\; \cst{g}\;y\;x$
as the witness for~$f$.

### 2022-03-04 pre/pre.tex
solid
more

\section{First-Order Logic}

@@

\subsection{Term Order and Selection}
\label{sec:pre:order}
Superposition calculus is parametrized by a {\em reduction ordering} $\succ$, an
ordering on terms that has the following properties:

>>>>

\section{First-Order Logic}
\label{sec:pre:fol}

@@

\subsection{Term Order and Selection}
\label{sec:pre:order}
Superposition calculus is parameterized by a {\em reduction ordering} $\succ$, an
ordering on terms that has the following properties:

### 2022-03-04 unif/unif.tex
more

The transitions are parametrized by a mapping~$\mathcal{P}$ that assigns a set

@@

the transitions are parametrized by a selection function~$S$ mapping a multiset

@@

Like the JP procedure, our procedure, parametrized with $\mathcal{P}_\mathsf{c}$,  misses no unifiers:

\begin{theorem}
\label{thm:completeness}
% \looseness=-1
The procedure described in Section \ref{sec:unif:the-unification-procedure}
parametrized by $\mathcal{P}_\mathsf{c}$ is complete, meaning that the
substitutions on the leaves of the constructed tree form a CSU.
More precisely, let $E$ be a multiset of constraints
and let $V$ be the supply of fresh variables provided to the procedure. 
Then for any unifier $\varrho$ of $E$ there exists
a derivation $\left(E,\makeop{id}\right)\longrightarrow^{*}\sigma$ and a
substitution $\theta$ such that for all free variables $X\not\in V$, we have $\varrho X = \theta\sigma X$.
\end{theorem}

@@

\begin{lemma}[$\jp L{3\text{ (1)}}$]\label{lem:principal-lemma}Let $\varrho$ be a substitution
    and $X$, $Y$ be free variables such that $\varrho\left(X\,\seq s\right)=\varrho\left(Y\,\seq t\right)$
    for some term tuples $\seq s$ and $\seq t$. Then for every opponent
    pair $u$, $v$ in $\varrho X$ and $\varrho Y$
    (Definition \ref{unif:def:opponents}), the head of $u$ or $v$ is a parameter
    of $\varrho X$ or $\varrho Y$.
    \end{lemma}
  %
\begin{exa}
  Let $\varrho$ be the substitution $\{ X \mapsto \lamx{\cst{f} \, x}, Y
  \mapsto \lamx{x} \}$. Clearly, $\varrho$ unifies the problem $X \, \cst{a}
  \unif Y \, (\cst{f} \, \cst{a})$. The opponent pair for $\varrho\,X$ and $\varrho\,Y$
  is $(\cst{f} \, x, y)$ which clearly adheres to the conclusion of Lemma \ref{lem:principal-lemma}.
%
  Intuitively, one of the consequences of this lemma is that any solution for
  flex-flex problems that does not mention parameters will not have opponent pairs. 
  For example, one solution for the 
  problem $X \, \cst{a}
  \unif Y \, (\cst{f} \, \cst{a})$ is $\varrho = \{ X \mapsto \cst{c}, Y \mapsto \cst{c} \}$
  which has no opponent pairs.
\end{exa}

  In contrast to applied constants, applied variables should not be eagerly decomposed.
  For a constant $\cst{f}$, if $\cst{f}\,\seq s \unif \cst{f}\,\seq t$ has a unifier,
  that unifier must clearly also unify $s_{i} \unif t_{i}$ for each~$i$. 
  For a free variable $X$, a unifier of $X\,\seq s \unif X\,\seq t$ does not necessarily
  unify $s_{i} \unif t_{i}$. For example, $\{X \mapsto \cst{a}\}$ is a unifier that does
  refer to any of the arguments of $X$.
  The concept of $\omega$-simplicity is
  a criterion on unifiers that captures some of the cases where eager decomposition is possible.
  Non-$\omega$-simplicity
  on the other hand is the main trigger of $\unifrulename{Iteration}$\textemdash the
  most explosive binding of our procedure.

@@

  To compare positions of $s$ and $s\,\seq t$, ignore
  the parameter count mismatch.
  % ASK(ALEX), ASK(VISA): I am no expert, but isn't it unusual to use these
  % kind of instructions in the statement of the theorem? Maybe it would be good
  % to locally redefine the concept of position.

@@

        \overline{u}_m$. Since PT is parametrized by an admissible

@@

\looseness=-1
We used OCaml's functors to create a modular implementation. The core of our
procedure is implemented in a module which is parametrized by another module
providing oracles and implementing the \textsf{Bind} step. In this way we can
obtain the complete or pragmatic procedure and seamlessly integrate
oracles while reusing as much common code as possible.

>>>>

The transitions are parameterized by a mapping~$\mathcal{P}$ that assigns a set

@@

the transitions are parameterized by a selection function~$S$ mapping a multiset

@@

Like the JP procedure, our procedure, parameterized with $\mathcal{P}_\mathsf{c}$,  misses no unifiers:

\begin{theorem}
\label{thm:completeness}
% \looseness=-1
The procedure described in Section \ref{sec:unif:the-unification-procedure}
parameterized by $\mathcal{P}_\mathsf{c}$ is complete, meaning that the
substitutions on the leaves of the constructed tree form a CSU.
More precisely, let $E$ be a multiset of constraints
and let $V$ be the supply of fresh variables provided to the procedure. 
Then for any unifier $\varrho$ of $E$ there exists
a derivation $\left(E,\makeop{id}\right)\longrightarrow^{*}\sigma$ and a
substitution $\theta$ such that for all free variables $X\not\in V$, we have $\varrho X = \theta\sigma X$.
\end{theorem}

@@

\begin{lemma}[$\jp L{3\text{ (1)}}$]\label{lem:principal-lemma}Let $\varrho$ be a substitution
    and $X$, $Y$ be free variables such that $\varrho(X\,\seq s)=\varrho(Y\,\seq t)$
    for some term tuples $\seq s$ and $\seq t$. Then for every opponent
    pair $u$, $v$ in $\varrho X$ and $\varrho Y$
    (Definition \ref{unif:def:opponents}), the head of $u$ or $v$ is a parameter
    of $\varrho X$ or $\varrho Y$.
    \end{lemma}
  %
\begin{exa}
  Let $\varrho$ be the substitution $\{ X \mapsto \lamx{\cst{f} \, x}, Y
  \mapsto \lamx{x} \}$. Clearly, $\varrho$ unifies the problem $X \, \cst{a}
  \unif Y \, (\cst{f} \, \cst{a})$. The opponent pair for $\varrho\,X$ and $\varrho\,Y$
  is $(\cst{f} \, x, y)$ which clearly adheres to the conclusion of Lemma \ref{lem:principal-lemma}.
%
  Intuitively, one of the consequences of this lemma is that any solution for
  flex-flex problems that does not mention parameters will not have opponent pairs. 
  For example, one solution for the 
  problem $X \, \cst{a}
  \unif Y \, (\cst{f} \, \cst{a})$ is $\varrho = \{ X \mapsto \lamx{\cst{c}}, Y \mapsto \lamx{\cst{c}} \}$
  which has no opponent pairs.
\end{exa}

  In contrast to applied constants, applied variables should not be eagerly decomposed.
  For a constant $\cst{f}$, if $\cst{f}\,\seq s \unif \cst{f}\,\seq t$ has a unifier,
  that unifier must clearly also unify $s_{i} \unif t_{i}$ for each~$i$. 
  For a free variable $X$, a unifier of $X\,\seq s \unif X\,\seq t$ does not necessarily
  unify $s_{i} \unif t_{i}$. For example, $\{X \mapsto \lam{\overline{x}}{\cst{a}}\}$ is a unifier that does
  refer to any of the arguments of $X$.
  The concept of $\omega$-simplicity is
  a criterion on unifiers that captures some of the cases where eager decomposition is possible.
  Non-$\omega$-simplicity
  on the other hand is the main trigger of $\unifrulename{Iteration}$\textemdash the
  most explosive binding of our procedure.

@@

  To compare positions of $s$ and $s\,\seq t$, we ignore
  the outermost $\lambda$-binders of s.

@@

        \overline{u}_m$. Since PT is parameterized by an admissible

@@

\looseness=-1
We used OCaml's functors to create a modular implementation. The core of our
procedure is implemented in a module which is parameterized by another module
providing oracles and implementing the \textsf{Bind} step. In this way we can
obtain the complete or pragmatic procedure and seamlessly integrate
oracles while reusing as much common code as possible.
